>

### 2024-04-02 docs/article.qmd
Draft initial intro and methods

---
title: >
  Eigenvector centrality as a metric of polypharmacy signifies antidepressants
  as a central hub of the psychopharmaca co-prescription network

format:
  elsevier-pdf:
    latex-output-dir: _render
    include-in-header:
      file: _header.tex
    journal:
      formatting: review
      cite-style: authoryear

funding: "The author(s) received no specific funding for this work."
keywords:
  - Graph Theory
  - Prescription Registry
  - Polypharmacy
  - Mental Health   
  - Public Health Monitoring

@@

In psychiatry, polypharmacy is defined as a patient with at least two concurrent psychiatric medications [@Shrivastava2013]. With a prevalence between 13-90%, psychiatric polypharmacy manifests in five categories, namely same-class, multi-class, adjunctive, augmentation, and total polypharmacy. Same-class polypharmacy is the use of multiple medications from the same class. Multi-class polypharmacy is the use of multiple medications from different classes indicated for the same symptom cluster. Adjunctive polypharmacy is the use of additional medications to treat side effects due to other medications. Augmentation polypharmacy is the use of full-dose and sub-dose medications from a different class for the same symptom cluster. Finally, total polypharmacy is the total count of all medications used by a patient, both psychopharmaca and other classes alike.

Paragraph layout:

- Mental disorders as a common co-morbidity contributes to polypharmacy in general
- Specific types of polypharmacy in mental disorders
- Problem: The need of a systematic public health monitoring for polypharmacy
- Definition and purpose of the study:
  - Proposed solution: Drug prescription network and network analysis to evaluate polypharmacy
  - Question: How do we implement network analysis approach to quantitatively identify polypharmacy?
  - Objective: Use graph centrality measure as a quantitative indicator of polypharmacy to enable systematic public health monitoring

- No agreement on a quantitative indicator of polypharmacy
- Problem of exposure: Polypharmacy can be defined as cumulative or simultaneous prescriptions

- Create a co-prescription network
- The weight is the number of time a pair of medications is prescribed together
- Drugs are classified per ATC code using five different levels
- The co-prescription network has a scale-free property, suggesting that the probability new links being formed is growing alongside the degree of the target node
- The scale-free property of the network might be explained by the physician's prescribing habit: The physician tends to prescribe medications with known medications

@@

## Exploratory descriptive analysis

@@

## Exploring seasonality in the dataset

@@

## Decomposition with singular spectrum analysis

@@

@@

>>>>

---
title: >
  Eigenvector centrality as a metric of polypharmacy signifies antidepressants
  as a central hub of the drug prescription network

format:
  elsevier-pdf:
    include-in-header:
      file: _header.tex
    journal:
      formatting: review
      cite-style: authoryear

funding: "The author(s) received no specific funding for this work."
keywords:
  - Graph Theory
  - Prescription Registry
  - Polypharmacy
  - Mental Health   
  - Public Health Surveillance

@@

In psychiatry, polypharmacy is defined as a patient with at least two concurrent psychiatric medications [@Shrivastava2013]. With a prevalence between 13-90%, psychiatric polypharmacy manifests in five categories, namely same-class, multi-class, adjunctive, augmentation, and total polypharmacy. Same-class polypharmacy is the use of multiple medications from the same class. Multi-class polypharmacy is the use of multiple medications from different classes indicated for the same symptom cluster. Adjunctive polypharmacy is the use of additional medications to treat side effects due to other medications. Augmentation polypharmacy is the use of full-dose and sub-dose medications from a different class for the same symptom cluster. Finally, total polypharmacy is the total count of all medications used by a patient, both psychopharmaca and other classes alike [@Shrivastava2013]. For policymakers, differences in defining polypharmacy pose a challenge to design a surveillance system [@Sirois2016].

Surveilling polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Stassen2021], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy surveillance system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. To reduce the complexity, graph theory offers a complementary approach by analyzing polypharmacy as a drug prescription network (DPN). Transforming prescription registry into a DPN has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

A DPN is constructed as a matrix which rows to columns represent pairs of medication being taken by a patient in a day. Since there is no limit for the number of rows and columns, a DPN can contain virtually every possible combination of polypharmacy regimens. A DPN is generalizable as a population aggregate by coalescing all single-patient DPN through matrix addition [@Cavallo2012]. Constructing a DPN allows examining co-prescription dynamics in the population as graph measures, which is also attributable to each medication. One of such measures is graph centrality, where we calculate scores representing the number of co-medications, number of connecting paths, average distance, and relative importance [@Askar2021]. The relative importance score is of particular interest to address concerns raised by @Sirois2016 and @Delara2022 regarding standardized quantitative indicator of polypharmacy. Relative importance, most commonly measured by eigenvector centrality, represents the connection to vital medications. A high eigenvector centrality in a DPN indicates more prescriptions with highly-prescribed medications. Following up @Cavallo2012 and @Bazzoni2015, we hypothesize that eigenvector centrality as a measure of relative importance is attributable to indicate polypharmacy. Our investigation showcases eigenvector centrality as a quantitative indicator of polypharmacy to promote a systematic public health surveillance system.

::: {.content-hidden}
> **"How do we implement a network analysis approach to quantitatively attribute polypharmacy?"**
:::

@@

## Data analysis

### Exploratory descriptive analysis

@@

### Exploring seasonality in the dataset

@@

### Decomposition with singular spectrum analysis

@@

- Problem of exposure: Polypharmacy can be defined as cumulative or simultaneous prescriptions
- Cut-off threshold of polypharmacy should be justified by the risk of deleterious health outcomes
- An ideal surveillance system should describe the importance of polypharmacy

@@

- The weight is the number of time a pair of medications is prescribed together
- The co-prescription network has a scale-free property, suggesting that the probability new links being formed is growing alongside the degree of the target node
- The scale-free property of the network might be explained by the physician's prescribing habit: The physician tends to prescribe medications with known medications

### 2024-04-02 docs/exec-summary/_preprocessing.qmd
Draft an initial report using Elsevier template
Draft initial intro and methods

### Data pre-processing

Daily medication claims is a tabular data containing values from single medication entity for each patients in the cohort. Data transformation was performed to construct a square matrix of aggregate medication claim per day, with rows and columns representing groups of medication. Each element in the upper and lower triangle represents co-prescriptions, while the diagonal of the matrix represents single prescriptions. All elements in the matrix is a tally of drug combination/single prescription in a population level claimed on a daily basis. The matrix follows a knowledge graph construct with a subject-predicate-object (s, p, o) triples, i.e. `Medication1-PrescribedAlongside-Medication2`. The knowledge graph is a representation of medication-claim network, where the edge (predicate) is the number of concurrency. Weight was applied on the edges by dividing it by the defined daily dose (DDD). Weighting is necessary to regularize the matrix into a unit-free scale, which allow a direct comparison of one drug to another.

>>>>

::: {.content-visible unless-meta="in.article"}
### Data pre-processing
:::

Daily medication claims is a tabular data containing values from single medication entity for each patients in the cohort. Data transformation was performed to construct a square matrix of aggregate medication claim per day, with rows and columns representing groups of medication. Each element in the upper and lower triangle represents co-prescriptions, while the diagonal of the matrix represents single prescriptions. All elements in the matrix is a tally of drug combination/single prescription in a population level claimed on a daily basis. The matrix follows a knowledge graph construct with a subject-predicate-object (s, p, o) triples, i.e. `Medication1-PrescribedAlongside-Medication2`. The knowledge graph is a representation of drug prescription network (DPN), where the edge (predicate) is the number of concurrency. In the original DPN, the edge weight was the number of claims by assuming that all of defined daily doses (DDD) were equal to 1 [@Cavallo2012]. To account for augmentation polypharmacy, DDD provides further information due to the presence of a sub-dose medication, where DDD < 1 should be allowed. Naturally, an augmentation polypharmacy regimen should have a higher weight since this regimen is intended to increase the efficacy of the augmented medication [@Shrivastava2013]. Therefore, weight was re-applied on the edges by dividing the number of claims by the DDD.

### 2024-04-02 docs/exec-summary/_source-data.qmd
Draft an initial report using Elsevier template
Draft initial intro and methods

### Source of data

The University of Groningen IADB.nl pharmacy prescription database is a growing database that contains prescription data for 31 years from 1994 till 2024 from approximately 120 community pharmacies and covers an estimated population of 1,120,000 patients. Registration in the database is irrespective of health care insurance and age, gender and prescription rates among the database population have been found to be representative of the Netherlands as a whole, and the database has been widely used for research. Each person is individually tracked throughout the database period and prescription records contain information on the date of dispensing, the quantity dispensed, the dose regimen, the number of days the prescription is valid, the prescribing physician and the Anatomical Therapeutic Chemical code (ATC code). Each patient has a unique anonymous identifier; date of birth and gender are known. Due to the high patient-pharmacy commitment in the Netherlands, the medication records for each patient are virtually complete, except for over the counter (OTC) drugs and medication dispensed during hospitalization. This analysis include daily medication claims from a static cohort of adults aged 18-65 years old and prescribed for anxiolytics or antidepressants at least once in the time span from 2018 to 2022.

>>>>

::: {.content-visible unless-meta="in.article"}
### Source of data
:::

All data in this investigation originated from IADB.nl, a dynamic pharmacy database containing prescription data from 1994 which still grow expansively. The database covers approximately 120 community pharmacies serving over one million patients. All patients are recorded in the database, irrespective of health care insurance, where the prescription rates, age, and gender are generalizable to represent the Netherlands [@Visser2013]. The database offers a longitudinal record, where each individual has a uniquely anonymized tracking identifier. All records contain fields on the date of medication dispensing, its quantity, dose regimen, number of days prescribed, prescribing physicians, and the Anatomical Therapeutic Chemical (ATC) code. Except for the over the counter drugs and medication dispensed during hospitalization, other medication records for each patient are complete, mainly due to high patient-pharmacy commitment in the Netherlands. This analysis include daily medication claims from a static cohort of adults aged 18-65 years old and prescribed for anxiolytics or antidepressants at least once in the time span from 2018 to 2022.
